|

### 2022-04-19 paper/paper.Rmd
Comment added
added section

---
title: |
  A new simplified manual tour, with examples in mathematica and R
type: Short Technical Note
author:
  - name: Alex Aumann
    affil: a
    email: aaum0002@student.monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
  - name: Ursula Laa
    affil: b
    email: ursula.laa@boku.ac.at
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Something here
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
output: rticles::tf_article
keep_tex: true
---

@@

@@

>>>>

---
title: |
  A new simplified manual tour, with examples in mathematica and R
type: Short Technical Note
author:
  - name: Alex Aumann
    affil: a
    email: aaum0002@student.monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
  - name: Ursula Laa
    affil: b
    email: ursula.laa@boku.ac.at
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Something here
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

### Algorithm 3

For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$. Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$. Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalised), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
\end{enumerate}
\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
\end{enumerate}
The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.

\begin{align}
A=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{P_1}{\downarrow} & \stackrel{P_2}{\downarrow} &  & &\\
e_1\to & x_{11} & x_{12}  & x_{13} & \cdots & x_{1p} \\
e_2 \to& x_{21}  & x_{22}  & \cdot  &\cdots & x_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& x_{m1}  & x_{m2}  & \cdot  &\cdots & x_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&x_{p1} & x_{p2} &\cdot &\cdots & x_{pp} 
\end{pNiceArray},\quad (x_{m1},x_{m2})\to (a,b)
\end{align}

@@

XXX need to check calculation of angle when $m_1 < 0$
